Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Paris version \YellowPaperVersionNumber}}

### 2023-12-21 Paper.tex
grammar edit
address comments round 2

\begin{description}
\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
\item[ommersHash] A 256-bit hash field that is now deprecated due the replacement of proof of work consensus. It is now to a constant, $\texttt{KEC}(\texttt{RLP}(()))$; formally $H_{\mathrm{o}}$.
\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which priority fees from this block are transferred; formally $H_{\mathrm{c}}$.
\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
\item[difficulty] A scalar field that is now deprecated due to the replacement of proof of work consensus. It is set to 0; formally $H_{\mathrm{d}}$.
\item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally \hyperlink{block_number_H__i}{$H_{\mathrm{i}}$}.
\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
\item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
\item[prevRandao]\linkdest{prevRandao_H__a}{}\linkdest{H__a} the latest RANDAO mix\footnote{RANDAO is a pseudorandom value generated by validators on the Ethereum consensus layer. Refer to the consensus layer specs (\url{https://github.com/ethereum/consensus-specs}) for more detail on RANDAO.} of the post beacon state of the previous block; formally $H_{\mathrm{a}}$.
\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value that is now deprecated due to the replacment of proof of work consensus. It is set to \texttt{0x0000000000000000}; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.
\item[baseFeePerGas]\linkdest{block_baseFeePerGas_H__f}{} A scalar value equal to the amount of wei that is burned for each unit of gas consumed; formally \hyperlink{H__f}{$H_{\mathrm{f}}$}.
\end{description}
\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are a series of transactions, $B_{\mathbf{T}}$, and an empty array which was previously reserved for ommer block headers, $B_{\mathbf{U}}$. Formally, we can refer to a block $B$:
\begin{equation}
B \equiv (B_{\mathrm{H}}, B_{\mathbf{T}}, B_{\mathbf{U}})
\end{equation}

@@

\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: its ommers field $B_{\mathbf{U}}$ must be an empty array and it must be internally consistent with the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
\begin{equation}
\begin{array}[t]{lclc}
B_{\mathbf{U}} &\equiv& [] & \wedge \\
\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p_{\mathrm{T}}(i, B_{\mathbf{T}}[i])\}) & \wedge \\
\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p_{\mathrm{R}}(i, B_{\mathbf{R}}[i])\}) & \wedge \\
\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
\end{array}
\end{equation}
where $p_{\mathrm{T}}(k, v)$ and $p_{\mathrm{R}}(k, v)$ are pairwise RLP transformations, but with a special treatment for EIP-2718 transactions:
\begin{equation}
p_{\mathrm{T}}(k, T) \equiv \left( \mathtt{RLP}(k), \begin{cases}
\mathtt{RLP}(\hyperlink{L_transaction}{L_{\mathrm{T}}}(T)) & \text{if} \quad T_{\mathrm{x}} = 0 \\
(T_{\mathrm{x}}) \cdot \mathtt{RLP}(L_{\mathrm{T}}(T)) & \text{otherwise}
\end{cases}
\right)
\end{equation}
and
\begin{equation}
p_{\mathrm{R}}(k, R) \equiv \left( \mathtt{RLP}(k), \begin{cases}
\mathtt{RLP}(\hyperlink{L__R}{L_{\mathrm{R}}}(R)) & \text{if} \quad R_{\mathrm{x}} = 0 \\
(R_{\mathrm{x}}) \cdot \mathtt{RLP}(L_{\mathrm{R}}(R)) & \text{otherwise}
\end{cases}
\right)
\end{equation}
($\cdot$ is the concatenation of byte arrays).

>>>>

\begin{description}
\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
\item[ommersHash] A 256-bit hash field that is now deprecated due to the replacement of proof of work consensus. It is now to a constant, $\texttt{KEC}(\texttt{RLP}(()))$; formally $H_{\mathrm{o}}$.
\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which priority fees from this block are transferred; formally $H_{\mathrm{c}}$.
\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
\item[difficulty] A scalar field that is now deprecated due to the replacement of proof of work consensus. It is set to 0; formally $H_{\mathrm{d}}$.
\item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally \hyperlink{block_number_H__i}{$H_{\mathrm{i}}$}.
\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
\item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
\item[prevRandao]\linkdest{prevRandao_H__a}{}\linkdest{H__a} the latest RANDAO mix\footnote{RANDAO is a pseudorandom value generated by validators on the Ethereum consensus layer. Refer to the consensus layer specs (\url{https://github.com/ethereum/consensus-specs}) for more detail on RANDAO.} of the post beacon state of the previous block; formally $H_{\mathrm{a}}$.
\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value that is now deprecated due to the replacment of proof of work consensus. It is set to \texttt{0x0000000000000000}; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.
\item[baseFeePerGas]\linkdest{block_baseFeePerGas_H__f}{} A scalar value equal to the amount of wei that is burned for each unit of gas consumed; formally \hyperlink{H__f}{$H_{\mathrm{f}}$}.
\end{description}
\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are a series of transactions, $B_{\mathbf{T}}$, and an empty array which was previously reserved for ommer block headers, $B_{\mathbf{U}}$. Formally, we can refer to a block $B$:
\begin{equation}
B \equiv (B_{\mathrm{H}}, B_{\mathbf{T}}, B_{\mathbf{U}})
\end{equation}

@@

\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: the block's ommers field $B_{\mathbf{U}}$ must be an empty array and the block's header must be consistent with the given transactions $B_{\mathbf{T}}$. For the header to be consistent with the transactions $B_{\mathbf{T}}$, \textbf{stateRoot} ($H_{\mathrm{r}}$) must match the resultant state after executing all transactions in order on the base state $\boldsymbol{\sigma}$ (as specified in section \ref{ch:finalisation}), and \textbf{transactionsRoot} ($H_{\mathrm{t}}$), \textbf{receiptsRoot} ($H_{\mathrm{e}}$), and \textbf{logsBloom} ($H_{\mathrm{b}}$) must be correctly derived from the transactions themselves, the transaction receipts resulting from execution, and the resulting logs, respectively.
\begin{equation}
\begin{array}[t]{lclc}
B_{\mathbf{U}} &\equiv& () & \wedge \\
\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p_{\mathrm{T}}(i, B_{\mathbf{T}}[i])\}) & \wedge \\
\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p_{\mathrm{R}}(i, B_{\mathbf{R}}[i])\}) & \wedge \\
\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
\end{array}
\end{equation}
where $p_{\mathrm{T}}(k, v)$ and $p_{\mathrm{R}}(k, v)$ are pairwise RLP transformations, but with a special treatment for EIP-2718 transactions:
\begin{equation}
p_{\mathrm{T}}(k, T) \equiv \left( \mathtt{RLP}(k), \begin{cases}
\mathtt{RLP}(\hyperlink{L_transaction}{L_{\mathrm{T}}}(T)) & \text{if} \quad T_{\mathrm{x}} = 0 \\
(T_{\mathrm{x}}) \cdot \mathtt{RLP}(L_{\mathrm{T}}(T)) & \text{otherwise}
\end{cases}
\right)
\end{equation}
and
\begin{equation}
p_{\mathrm{R}}(k, R) \equiv \left( \mathtt{RLP}(k), \begin{cases}
\mathtt{RLP}(\hyperlink{L__R}{L_{\mathrm{R}}}(R)) & \text{if} \quad R_{\mathrm{x}} = 0 \\
(R_{\mathrm{x}}) \cdot \mathtt{RLP}(L_{\mathrm{R}}(R)) & \text{otherwise}
\end{cases}
\right)
\end{equation}
($\cdot$ is the concatenation of byte arrays).
